% This starts the document; the "npsreport" style is really a modified
% report style. Feel free to use the options:
%
% "twoside" - prints "THIS PAGE INTENTIONALLY LEFT BLANK" on
%             odd-numbered pages at the end of sections.
%
% "traditional" - prints a ``traditional'' NPS thesis style, rather
% than the new one.
%
% See also /opt/local/share/texmf-dist/tex/latex/base/report.cls

\documentclass[twoside,thesis,authorindex]{npsreport} 
% Options include:
% twoside - always use at NPS
% thesis  - specify for a thesis
% authorindex - generate the author index (requires post-processing )
% traditional - make an ugly thesis

%
% Reference commands
%

% And here is the main thesis

\usepackage{amstext} %added by JHG to support \text in an equation
\usepackage{amsmath} %added by JHG to support more equation options
\usepackage{rotating} %add by JHG to allow sideways table labels
\usepackage{etex} %added by JHG to fix too many floats issue with tables in appendices
\reserveinserts{18}
\usepackage{morefloats}
\usepackage{hyperref}


\securitybanner{}
\title{Author Detection on a Mobile Phone}
\author{Jody Grady}
\degree{Master of Science in Computer Science}
\degreeabbreviation{MS}
\department{Department of Computer Science}
\thesisadvisor{Dr. Rob Beverly}
\secondreader{Dr. Craig Martell}
\departmentchair{Dr. Peter Denning}
\rank{Commander}
\prevdegrees{Bachelor of Engineering in Aerospace Engineering}
\degreedate{March 25, 2011} 
\distribution{Approved for public release; distribution is unlimited}
\RPTpreparedFor{}
\abstract{
Traditional author detection is conducted on powerful computers using documents such as books and articles.  With the explosion of mobile phone computing use, modern author detection needs to be lean enough to operate on a resource restrained mobile phone and robust enough to handle the terse and non-standard wording in text messages, Tweets, and emails.  By testing natural language and machine learning techniques for size and speed, not just effectiveness, this thesis identifies feature and technique combinations appropriate for author detection on a mobile phone.  Specifically this thesis will examine effectiveness versus storage size for word grams of size 1, 2, and 5 as well as Gappy Bigrams and Orthogonal Sparse Bigrams.  To deal with the robust nature of Tweets and text message, the Google Web1T corpus will be tested for size versus effectiveness in combination with the word grams.  Once appropriate feature and technique combinations are found, those combinations will be tested on actual Android mobile phones to gauge how effective the chosen techniques are on a real mobile phone.
}

% Information for the SF298.
% These are mandatory:
\ReportType{Master's Thesis}
\DatesCovered{2009-03-01 - 2011-03-25}
\SponsoringAgency{Department of the Navy}
\ReportClassification{Unclassified}
\AbstractClassification{Unclassified}
\PageClassification{Unclassified}

\SupplementaryNotes{ The views expressed in this thesis are those of
  the author and do not reflect the official policy or position of the
  Department of Defense or the U.S. Government. \textbf{\footnotesize
    IRB Protocol Number: XXXX}}

% These are optional:
%\ContractNumber{}
%\GrantNumber{}
%\ProgramElementNumber{}
%\TaskNumber{}
%\WorkUnitNumber{}
%\POReportNumber{}
%\Acronyms{}
%\SMReportNumber{}
%\SubjectTerms{}
%\ResponsiblePerson{}
%\RPTelephone{}

%
% your thesis begins here
%


% Prevent footnotes from being reset at each chapter
% Comment this out to have the reset with each chapter.
\makeatletter
\@removefromreset{footnote}{chapter}
\makeatother



% This must be on a line by itself:
\begin{document}

\NPScover                       % Cover
\NPSsftne                       % SF298
\NPSthesistitle                 % Title page
\NPSabstractpage                % Abstract Page
\NPSfrontmatter                 % NPS front matter follows

% This changes the chaptermark and includes the various tables
% It must be here

\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\chaptername}\ \thechapter.\ #1}{}}
\NPSTableOfContents
\NPSListOfFigures
\NPSListOfTables

% Put the acknowledgements here
\acknowledgements{
  This thesis would not have been possible without the guidance and instruction of Dr. Rob Beverly and Dr. Craig Martell.  Also, Dylan Freedman, who interned at NPS for the summer of 2010, did tremendous work creating the minimum perfect hash files, signature files, and scripts to create even more hash files for this thesis.  His ability to grasp and implement complex hashes over a huge corpus of Google Web1T words was invaluable.  Of course, the patience show by my wife Kerri-Leigh and sons, Rowan and Aiden, was a major source of support for me while creating this thesis.
}


% NPS body follows
\NPSbody

%\chapter{[CHAPTER TITLE]}
%This is the beginning of your thesis. Don't be a Micky Mouse\cite{mm2}. 

%[CHAPTER BODY]
\include{chapter1}
\include{chapter2}
\include{chapter3}
\include{chapter4}
\include{chapter5}


%\section{Your First Section}
%[Section Body]
%\section{Your Second Section}
%[Section Body]

\def\showURL{}
\bibliographystyle{nps-plain-unsorted}
\bibliography{thesis}

\NPSappendix                   % Appendices follow
%\chapter{[APPENDIX NAME]}
%[APPENDIX BODY]

%\include{appendixDummy}
\include{appendixA} %LibLinear Accuracy for ENRON
\include{appendixB} %LibLinear Accuracy for Twitter
\include{appendixC} %Naive Bayes Accuracy for ENRON
\include{appendixD} %Naive Bayes Accuracy for Twitter
\include{appendixE} %Grouped LibLinear Accuracy for ENRON
\include{appendixF} %Grouped LibLinear Accuracy for Twitter
\include{appendixG} %Grouped Naive Bayes Accuracy for ENRON
\include{appendixH} %Grouped Naive Bayes Accuracy for Twitter
\include{appendixI} %Liblinear Score ( Accuracy/Size ) for ENRON
\include{appendixJ} %Liblinear Score ( Accuracy/Size ) for Twitter
\include{appendixK} %Naive Bayes Score ( Accuracy/Size ) for ENRON
\include{appendixL} %Naive Bayes Score ( Accuracy/Size ) for Twitter
\include{appendixM} %Liblinear Sizes for ENRON
\include{appendixN} %Liblinear Sizes for Twitter
\include{appendixO} %Naive Bayes Sizes for ENRON
\include{appendixP} %Naive Bayes Sizes for Twitter
%\include{appendixQ} %Liblinear Enron Plot of Grouping Behavior
%\include{appendixR} %Liblinear Enron Plot of Grouping Behavior
%\include{appendixS} %Liblinear Enron Plot of Grouping Behavior
%\include{appendixT} %Liblinear Enron Plot of Grouping Behavior
%\include{appendixU} %Liblinear Enron Plot of Grouping Behavior


% This is the official end of the thesis
\NPSend         

% This is the official distribution list.
% Free free to comment out or add your own
% The list is automatically properly numbered.

\chapter*{Initial Distribution List}
\addcontentsline{toc}{chapter}{Initial Distribution List}
\singlespace
\begin{enumerate}
\item Defense Technical Information Center\\Ft. Belvoir, Virginia
\item Dudly Knox Library\\Naval Postgraduate School\\Monterey, California

\end{enumerate}


\end{document}

% LocalWords:  

